---
title: "Possums are Awesome"
author: "April & Yi"
date: "Winter 2019"
output: 
  ioslides_presentation: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Set Background
The possum data frame consists of nine morphometric measurements on 104 mountain brushtail possums, trapped at seven sites from Southern Victoria to central Queensland. In 2002, they were proposed to be reclassified into two separate species: the mountain brushtail and short-eared.  

Source
> Lindenmayer, D. B., Viggers, K. L., Cunningham, R. B., and Donnelly, C. F. 1995. Morphological variation among columns of the mountain brushtail possum, Trichosurus caninus Ogilby (Phalangeridae: Marsupiala). Australian Journal of Zoology 43: 449-458.

> URL: https://vincentarelbundock.github.io/Rdatasets/datasets.html

## Load libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(skimr)
library("RColorBrewer")
```

## Read data
```{r message=FALSE, warning=FALSE}
 possum <- 
   readr::read_csv(file = "possum.csv", 
   na = c("", " ", "NA", "#N/A", "-999", "\\"))
```

```{r}
# Check dimensions
dim(possum)

# Check for NA's
##install.packages("skimr")
possum %>% 
  skimr::skim()

# Check column names
colnames(possum)
```

## Renaming variables for consistency
```{r}
possum <-
  possum%>% 
  dplyr::rename(
    pop = Pop, #population
    foot = footlgth, #foot length
    head = hdlngth, #head length
    skull = skullw, #skull width
    tail = taill, #tail length
    tooth = totlngth #tooth length
    ) 
possum
```

## Dataset Cleaning-Up

1. Is the data frame tidy? Why or why not?

This data frame is not tidy, because the first column does not have a column name (it got assigned X1 as its name automatically by R).

How to change that:

```{r}
colnames(possum)<-c("sample", 
                    "case", 
                    "site", 
                    "pop", 
                    "sex",
                    "age",
                    "head",
                    "skull", 
                    "tooth", 
                    "tail", 
                    "foot", 
                    "earconch", 
                    "eye", 
                    "chest", 
                    "belly")
```

## NAs

1. Identify the NAs.

1) Which column has the most NAs? 
```{r}
possum %>%   
  purrr::map_df(~ sum(is.na(.))) %>% 
  tidyr::gather(key="variables", value="num_nas") %>% 
  arrange(desc(num_nas))
```



# Potential questions to ask
**Main question 1**
Is there a difference in possum morphology between populations in Victoria and populations outside of Victoria?
(to see if this data can justify seperating the possum into two species by mophometrics)

--> There's a greater difference in earconch and foot length between the two populations.

```{r}
possum %>% 
  gather(head, skull, tooth, tail, foot, earconch, eye, chest, belly, key=morph, value=measure) %>%
  group_by(pop) %>% 
  ggplot(aes(x=morph, y=measure, fill=pop))+
    geom_boxplot()
```



- insert more background info on the 2002 research that wanted to reclassify it into 2 different species, more stuff on organisms and external environment, phenotypic variation blahblah, hypothesis on whether we think victorian possums would look a certain way because of ____ or something  :) -

**Sub questions**
1. Difference in ear conch length between both populations



2. Correlation between age and ear conch length (because ear conch may shrink as possums age?)

 --> No correlation between age and ear conch length. 
```{r}
possum %>% 
  ggplot(aes(x=age, y=earconch, color = pop))+
  geom_point()+
  labs(title = "Relationship Between Possum Age and Ear Conch Length",
       x= "Age",
       y= "Ear Conch Length")
```



3. Compare ear conch length by sex in both populations (in case different genders of possum have different ear conch size)

4. 

5. 

6. 


**Main Question Two**
Examine if it is true that the two populations of possum show little sexual dimorphism. 


**Sub questions**

1. Number of each gender in both populations.

```{r}
possum %>% 
  filter (pop == "Vic") %>% 
  ggplot(aes(x=sex))+
  geom_bar()
```

```{r}
possum %>% 
  filter (pop == "other") %>% 
  ggplot(aes(x=sex))+
  geom_bar()
```

2. 
```{r}
possum %>% 
  filter(pop == "Vic") %>% 
  gather(head, skull, tooth, tail, foot, earconch, eye, chest, belly, key=morph, value=measure) %>%
  group_by(sex) %>% 
  ggplot(aes(x=morph, y=measure, fill=sex))+
    geom_boxplot()+
  labs()
```

```{r}
possum %>% 
  filter(pop == "other") %>% 
  gather(head, skull, tooth, tail, foot, earconch, eye, chest, belly, key=morph, value=measure) %>%
  group_by(sex) %>% 
  ggplot(aes(x=morph, y=measure, fill=sex))+
    geom_boxplot()+
  labs()
```

# Conclusion 

(in the event that we can't find much of a correlation, we explain why we think so)


**Here's some outside research on the possums:**

*Australian museum*
https://australianmuseum.net.au/learn/animals/mammals/mammology-collection-mountain-brushtail-possum/
*Geographic dimorphism in the mountain brushtail possum (Trichosurus caninus): The case for a new species*
https://www.researchgate.net/publication/240510066_Geographic_dimorphism_in_the_mountain_brushtail_possum_Trichosurus_caninus_The_case_for_a_new_species
